---
title: "Association curves figures"
author: "Yanelli Nunez"
date: "1/28/2022"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:  

Plot the association curves for the main analysis

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
# replace with project directory
project <- "path/to/project"
project.folder <- paste0(project, "USA_emissions_code/")

# add folder locations
#source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
#source(paste0(file.locations.folder,'external_file_locations.R'))
#packages.folder =  paste0(project.folder, "code/packages/")
# load packages
source(paste0(project.folder,'code/packages/packages_to_load.R'))

code.folder = paste0(project.folder, "code/")
# load functions
source(paste0(code.folder, 'models/functions.R'))

# load data

source(paste0(code.folder, 'data_loading.R'))
```


## 1. Load datasets 

Loading in the model results. These models were produced through the 'y_01_single-exp_nested_models_ns_mean_ref.Rmd' from the models folder

```{r}
model_data <- paste0(project.folder, "code/models")
models_linear <- readRDS(paste0(model_data, "/models_linear_revised.rds")) %>%
  arrange(outcome)

models_nonlin_asian_an <- readRDS(paste0(model_data, "/models_nonlin_asian_an_revised.rds")) %>%
  arrange(outcome)

models_nonlin_blk_wht <- readRDS(paste0(model_data, "/models_nonlin_blk_wht_revised.rds")) %>%
  arrange(outcome)

models_nonlin_eco_hisp <- readRDS(paste0(model_data, "/models_nonlin_eco_hisp_revised.rds")) %>%
  arrange(outcome)

models_nonlin_wht <- readRDS(paste0(model_data, "/models_nonlin_wht_revised.rds")) %>%
  arrange(outcome)
```


### 1a. Combine datasets
Merging some of the datasets together to make the next steps easier

```{r}

models_nonlin_wht <- models_nonlin_wht %>%
  dplyr::rename(plots = plots_white)

models_nonlinear <- rbind(models_nonlin_asian_an, models_nonlin_eco_hisp) %>%
  rbind(., models_nonlin_wht)

rm(models_nonlin_asian_an)
rm(models_nonlin_eco_hisp)
rm(models_nonlin_wht)

```


## 2. Extract plots into environment

```{r}

### nonlinear
nonlinear_rnm <- models_nonlinear %>%
  unite(name, outcome:exposure, sep = "_") %>%
  mutate(name = str_remove(name, "_relat_chg"))

# extract plots to environment using assign and foreach()
foreach(i = iterators::icount(nrow(nonlinear_rnm)), # iterate across rows
                             .combine = rbind) %do% { 
  
assign(nonlinear_rnm$name[[i]], nonlinear_rnm$plots[[i]])
                             }

### linear 
linear_rnm <- models_linear %>%
   unite(name, outcome:exposure, sep = "_") %>%
  mutate(name = str_remove(name, "_relat_chg"))


# extract nonlinear using assign and foreach()
foreach(i = iterators::icount(nrow(linear_rnm)), # iterate across rows
                             .combine = rbind) %do% { 
  
assign(linear_rnm$name[[i]], linear_rnm$plots[[i]])
                             }


### other white and black nonlinear models
nh3_agric_perc_white <- models_nonlin_blk_wht$plots_white[[1]]
nox_commer_perc_white <- models_nonlin_blk_wht$plots_white[[2]]
nox_trans_perc_white <- models_nonlin_blk_wht$plots_white[[3]]
so2_energy_perc_white <- models_nonlin_blk_wht$plots_white[[4]]

nh3_agric_perc_black <- models_nonlin_blk_wht$plots[[1]]
nox_commer_perc_black <- models_nonlin_blk_wht$plots[[2]]
nox_trans_perc_black <- models_nonlin_blk_wht$plots[[3]]
so2_energy_perc_black <- models_nonlin_blk_wht$plots[[4]]

```


## 2. Plots Economy variables

### 2a. Histograms

Density plots for the independent variables

```{r}

dens_unempl <- data_master %>%
  ggplot(aes(x = perc_unemp)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Unemployment") 

dens_poverty <- data_master %>%
  ggplot(aes(x = perc_pvert)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Poverty")

dens_income <- data_master %>%
  ggplot(aes(x = fam_incom_adjs)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("Median Family Income ($1000)") +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))  

dens_propty <- data_master %>%
  ggplot(aes(x = propty_val_adjs)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("Median Propty Value ($1000)") +
 scale_x_continuous(labels=function(x)x/1000)  
  
```


### 2b. Adjust axis 

Adjust the scale of the x-axis for income and property value

```{r}
# fix income axis scale

so2_indus_fam_incom_adjs <- so2_indus_fam_incom_adjs +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))  

nox_energy_fam_incom_adjs <- nox_energy_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

so2_energy_fam_incom_adjs <- so2_energy_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

nh3_agric_fam_incom_adjs <- nh3_agric_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

nox_trans_fam_incom_adjs <- nox_trans_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

oc_resid_fam_incom_adjs <- oc_resid_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

nox_commer_fam_incom_adjs <- nox_commer_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

# fix property value axis scale

so2_indus_propty_val_adjs <- so2_indus_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

nox_energy_propty_val_adjs <- nox_energy_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

so2_energy_propty_val_adjs <- so2_energy_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

nh3_agric_propty_val_adjs <- nh3_agric_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

nox_trans_propty_val_adjs <- nox_trans_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

oc_resid_propty_val_adjs <- oc_resid_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000)

nox_commer_propty_val_adjs <- nox_commer_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000)
```


### 2b. Ensemble panel figure economic variables

The code chunk produces the manuscripts' Figure 3

Note the final figure was polished in power point

```{r}

pdf(file = "curves_eco.pdf", width = 14, height = 20)            
 grid.arrange(dens_unempl, 
              dens_poverty,
              dens_income,
              dens_propty,
   
              so2_indus_perc_unemp,
              so2_indus_perc_pvert,
              so2_indus_fam_incom_adjs,
              so2_indus_propty_val_adjs,
              
              so2_energy_perc_unemp,
              so2_energy_perc_pvert,
              so2_energy_fam_incom_adjs,
              so2_energy_propty_val_adjs,
              
              nox_energy_perc_unemp,
              nox_energy_perc_pvert,
              nox_energy_fam_incom_adjs,
              nox_energy_propty_val_adjs,
              
              nh3_agric_perc_unemp,
              nh3_agric_perc_pvert,
              nh3_agric_fam_incom_adjs,
              nh3_agric_propty_val_adjs,
              
              nox_trans_perc_unemp,
              nox_trans_perc_pvert,
              nox_trans_fam_incom_adjs,
              nox_trans_propty_val_adjs,
              
              oc_resid_perc_unemp,
              oc_resid_perc_pvert,
              oc_resid_fam_incom_adjs,
              oc_resid_propty_val_adjs,
              
              nox_commer_perc_unemp,
              nox_commer_perc_pvert,
              nox_commer_fam_incom_adjs,
              nox_commer_propty_val_adjs,
              
             nrow = 8, ncol = 4)
            # left = gridtext::richtext_grob("Industry: SO<sub>2</sub> (% chg)", rot=90,
             #gp=gpar(fontsize=18,font=8)))
dev.off() 

```

The following two plots were made for FMPH 491:

This plot charts the % Poverty and % Unemployment on the same plot.

Use the data extracted from the plots made in the models. This is done because 
the data for the plots is not saved anywhere. See line 72 in 
USA_emissions_code/code/Figures/functions.R to see how the data is calculated. 

```{r}
df_so2_ind = rbind(mutate(so2_indus_perc_unemp$data, type = "% Unemployment"), 
           mutate(so2_indus_perc_pvert$data, type = "% Poverty"))
plot_ind <- ggplot(df_so2_ind, aes(x = indp_var,  group = type, color = type)) +
    theme_minimal() +
    #geom_ribbon(aes(ymin = lci, ymax = uci), fill = "blue", alpha = 0.3) +
    geom_path(aes(y = beta), size = 1) +
    geom_hline(yintercept = 0, linetype = 2, size =0.5) +
    xlab("Percent (%)") +
    ylab(expression("Percent Change (%) Industry SO"["2"])) +
  annotate("text", x=54, y = 10, label = "% Poverty") + 
  annotate("text", x=32, y = 60, label = "% Unemployment") +
  annotate("text", x=64, y = 1, label = "No Change") +
  theme(legend.position="none") +
  ggtitle(expression("Association of Socioeconomic Groups with Industry SO"["2"]))
plot_ind
ggsave("socioeconomic.png")
```

This code make two charts: the percent change in Industry SO2 again vs. Property value and another vs. Familly income

```{r}
plot_eng_income <- ggplot(so2_indus_fam_incom_adjs$data, aes(x = indp_var/1000)) +
    theme_minimal() +
    geom_path(aes(y = beta), size = 1, color = "#1A85FF") +
    geom_hline(yintercept = 0, linetype = 2, size =0.5) +
    xlab("Family Income ($1000)") +
    ylab("") +
    ylim(c(-24,12)) +
  # remove the tick makrs because these plots are on the same scale and they number don't need to be repeted twice
    theme(axis.text.y = element_blank(), 
      axis.ticks.y = element_blank(), 
      axis.title.y = element_blank(), aspect.ratio = 1)
plot_eng_income

df = mutate(so2_indus_propty_val_adjs$data, indp_var = indp_var/1000)
plot_eng_property <- ggplot(df, aes(x = indp_var)) +
    theme_minimal() +
    geom_path(aes(y = beta), size = 1, color = "#D41159") +
    geom_hline(yintercept = 0, linetype = 2, size =0.5) +
    xlab("Property Value ($1000)") +
    ylim(c(-24,12)) +
  theme(aspect.ratio = 1) +
  ylab(expression("Percent Change (%) Industry SO"["2"])) +
  annotate("text", x=1100, y = 1, label = "No Change")
plot_eng_property

p = grid.arrange(plot_eng_property,plot_eng_income, 
                 nrow = 1, 
                 respect=T,
                 # include the title in the grid arrange
                 top=textGrob(expression("Association of Property Value and Family Income with Industry SO"["2"]), x = 0, hjust = 0))
ggsave("Income_property_plot.png", p)
```


## 2. Plots Race variables

Note that these figures where edited in *y_06_plots_race_modified_axis.Rmd* to address reviewers comments pertaining to visibility of the curve. Thus, the final manuscript figure 2 is NOT the one generated here but in the *y_06_plots_race_modified_axis.Rmd*. 

### 2a. Histograms

Density plots for economic variables

```{r, eval = F}

dens_black <- data_master %>%
  ggplot(aes(x = perc_black)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Black") +
  xlim(-3,86.5)

dens_white <- data_master %>%
  ggplot(aes(x = perc_white)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% White") +
  xlim(0, 104)

dens_asian <- data_master %>%
  ggplot(aes(x = perc_asian)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Asian") +
  xlim(-0.28,45)


dens_am_nat <- data_master %>%
  ggplot(aes(x = perc_nativ)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% American Native") +
  xlim(-0.17, 95)

dens_hisp <- data_master %>%
  ggplot(aes(x = perc_hisp_org)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Hispanic") +
  xlim(-0.65, 100)

```

### 2b. Edit axis of Hispanic

```{r, eval=F}

so2_indus_perc_hisp_org <- so2_indus_perc_hisp_org +
   xlim(0,100) 

so2_energy_perc_hisp_org <- so2_energy_perc_hisp_org +
  xlim(0,100)

nox_energy_perc_hisp_org <- nox_energy_perc_hisp_org +
  xlim(0,100)

nh3_agric_perc_hisp_org  <- nh3_agric_perc_hisp_org  +
xlim(0,100)

nox_trans_perc_hisp_org <- nox_trans_perc_hisp_org +
xlim(0,100)

oc_resid_perc_hisp_org <-  oc_resid_perc_hisp_org +
xlim(0,100)

nox_commer_perc_hisp_org <- nox_commer_perc_hisp_org +
xlim(0,100)

```


### 2c. Ensemble panel figure race

Produce manuscript's Figure 2
Note the final figures were polished in power point 

```{r, eval =F}

pdf(file = "curves_race.pdf", width = 14, height = 20)            
 grid.arrange(dens_white,
              dens_black,
              dens_asian,
              dens_am_nat,
              dens_hisp,
   
              so2_indus_perc_white,
              so2_indus_perc_black,
              so2_indus_perc_asian,
              so2_indus_perc_nativ,
              so2_indus_perc_hisp_org,
              
              so2_energy_perc_white,
              so2_energy_perc_black, 
              so2_energy_perc_asian,
              so2_energy_perc_nativ,
              so2_energy_perc_hisp_org,
              
              nox_energy_perc_white,
              nox_energy_perc_black,
              nox_energy_perc_asian,
              nox_energy_perc_nativ,
              nox_energy_perc_hisp_org,
              
              nh3_agric_perc_white,
              nh3_agric_perc_black, 
              nh3_agric_perc_asian,
              nh3_agric_perc_nativ,
              nh3_agric_perc_hisp_org,
              
              nox_trans_perc_white,
              nox_trans_perc_black, 
              nox_trans_perc_asian,
              nox_trans_perc_nativ,
              nox_trans_perc_hisp_org,
              
              oc_resid_perc_white,
              oc_resid_perc_black,
              oc_resid_perc_asian,
              oc_resid_perc_nativ,
              oc_resid_perc_hisp_org,
              
              nox_commer_perc_white,
              nox_commer_perc_black,  
              nox_commer_perc_asian,
              nox_commer_perc_nativ,
              nox_commer_perc_hisp_org,
              
             nrow = 8, ncol = 5)
            # left = gridtext::richtext_grob("Industry: SO<sub>2</sub> (% chg)", rot=90,
             #gp=gpar(fontsize=18,font=8)))
dev.off() 

```

